Analysing Constraint Grammar with a SAT solver

### 2015-10-23 paper/lrec/eval_futurework.tex
text

We ran for all three grammars the following thing.
For each rule, run all the rules before it to the symbolic sentence, and then ask if that rule can fire.
This test revealed problems in all grammars.
The Dutch grammar had the following rule sequence, making the latter impossible to apply:

\begin{itemize}
\item[] \texttt{REMOVE adv IF (1 n|np) ;}
\item[] \texttt{REMOVE adv IF (-1 det) (0 adj) (1 n|np) ;}
\end{itemize}

The Spanish grammar had an erroneous set definition, where a word was required to have almost all POS tags at the same time. After fixing the set definition, our test didn't reveal other problems.
The program marked __ problem rules for the Finnish grammar, but we didn't have time to verify if they were problematic or if our program is buggy.

As seen in table~\ref{table:res}, checking every rule in the Finnish grammar takes currently 3.5 hours.
A rule takes longer to run if it appears later in the rule sequence: 1000 previous rules is a larger SAT problem than 10 previous rules.
This can be seen from the time it took to compute the result for the last rule. For Finnish, the last rule took almost 4 minutes---had this been the average, the whole grammar would have taken 18 hours instead of 3.5.
We also measured the running times for the longest rules in the grammar.
\todo{find longest rule, change place of that rule, see what matters more. Promise: we can be smarter about SAT things.}

For future work, we want to improve the performance and scale up to larger grammars, and handle the full expressivity of CG-3, with \textsc{map}, \textsc{add} and \textsc{substitute} rules.
We want to detect more kinds of conflicts.
For instance, say we have the following rules:

@@

With our symbolic sentence, this rule set will be no problem; after all, there can be something else in the target beside noun and a verb.
However, if the grammar writer actually means the following

@@

then those rules are contradictory, and we would like to find it out.

\todo{Compare with a corpus-based test: do we find the same problems?}

>>>>

We tested all three grammars if they contain non-applicable rules:
for each rule, apply all the preceding rules to the symbolic sentence, and then ask if that last rule can fire.
This test revealed problems in all grammars.
The Dutch grammar had the following rules in the given order, making the latter impossible to apply:

\begin{itemize}
\item[] 
\begin{verbatim}REMOVE adv IF (1 n|np) ;
REMOVE adv IF (-1 det) (0 adj) (1 n|np) ;
\end{verbatim}
\end{itemize}

The Spanish grammar had an erroneous set definition, where a word was required to have almost all POS tags at the same time. After fixing the set definition, our test didn't reveal other problems.
The program marked \todo{n} problem rules for the Finnish grammar, but we didn't have time to verify if they were genuinely problematic or if our program is at fault.
We suspect that at least part of is related to the length of the sentences---for a condition that states $n$ or more words to the left or right, we only create a word exactly $n$ words away.

As seen in table~\ref{table:res}, checking every rule in the Finnish grammar takes currently 3.5 hours.
A rule takes longer to run if it appears later in the rule sequence: 1000 previous rules is a larger SAT problem than 10 previous rules.
This can be seen from the time it took to compute the result for the last rule. For Finnish, the last rule took almost 4 minutes---if this was the average speed, the whole grammar would have taken 18 hours instead of 3.5.
We also measured the running times for the longest rules in the grammar.
\todo{find longest rule, change place of that rule, see what matters more. Promise: we can be smarter about SAT things.}

For future work, we will continue improving the performance, and make sure that the implementation for the basic rules is correct; for instance, fixing the earlier mentioned issue with non-exact position.
After this, we want to handle the full expressivity of CG-3, with \textsc{map}, \textsc{add} and \textsc{substitute} rules.

Independently of 
As another line of investigation, we want to detect more kinds of conflicts.
For instance, say we have the following rules:

@@

With our symbolic sentence, these rules will be no problem; after all, there can be something else in the target beside noun and a verb.
However, it is likely that the grammar writer actually means the following:

@@

In that case, those rules are contradictory, and we would like to find it out.

Compare with a corpus-based test: do we find the same problems?
